Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing

### 2019-10-23 Dissonance_Santiago_v1.Rmd
Commit files from Paez-Laptop

---
title: Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Increasingly, however, it is recognized that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain users of public transport and active travel. A challenge with the way transportation policy is developed is that it tends to focus on cost-benefit and utilitarian considerations, which may miss other potential benefits of transportation. For this reason, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective and utilitarian values. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective and utilitarian values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associated to affective values (freedom, enjoyment, and happiness) and to utilitarian values (safety, time waste, and efficiency). The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with affective values.

@@

library(rticles)
library(readxl)
library(tidyverse)
library(rgdal)
library(sf)
library(ggmosaic)
library(ggthemes)
library(kableExtra)
```

@@

```{r figure-mosaic-plots-by-attribute-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-attribute}Mosaic plots for affective values; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value"}

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"))) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Mode associated with {Freedom, Enjoyment, Happiness, Poverty, Luxury, Status}") +
  ylab("Primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_wrap(~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r figure-mosaic-plots-by-attribute-and-age-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-age}Mosaic plots for affective values by age; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value"}

# Select only values that were significant in the tests of independence

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "LUXURY", "STATUS"))) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Mode associated with {Freedom, Enjoyment, Happiness, Luxury, Status}") +
  ylab("Primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_grid(AGE ~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r figure-mosaic-plots-by-attribute-and-education-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-education}Mosaic plots for affective values by education; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value"}

# Select only values that were significant in the tests of independence

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("LUXURY", "STATUS"))) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Mode associated with {Luxury, Status}") +
  ylab("Primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_grid(EDUCATION ~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r figure-mosaic-plots-by-attribute-and-income-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-income}Mosaic plots for affective values by income; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value"}

# Select only values that were significant in the tests of independence

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("POVERTY", "LUXURY", "STATUS"))) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Mode associated with {Poverty, Luxury, Status}") +
  ylab("Primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_grid(INCOME ~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r figure-mosaic-plots-by-attribute-and-time-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-travel-time}Mosaic plots for affective values by travel time; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value"}

# Select only affective values that were significant in the tests of independence

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("FREEDOM", "HAPPINESS", "LUXURY", "STATUS")) %>% drop_na(TRAVEL_TIME)) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Mode associated with {Freedom, Happiness, Luxury, Status}") +
  ylab("Primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_grid(TRAVEL_TIME ~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing
author:
  - name: Beasly Bee
    email: alice@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Antsy Aardvark
    email: aa@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Increasingly, however, it is recognized that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain users of public transport and active travel. A challenge with the way transportation policy is developed is that it tends to focus on cost-benefit and utilitarian considerations, which may miss other potential benefits of transportation. For this reason, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective and utilitarian values. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective and utilitarian values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associated to affective values (freedom, enjoyment, and happiness) and to utilitarian values (safety, time waste, and efficiency). The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with affective values.

@@

library(readxl)
library(tidyverse)
library(rgdal)
library(sf)
library(ggmosaic)
library(ggthemes)
library(kableExtra)
```

@@

```{r figure-mosaic-plots-by-attribute-without-instrumental, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-attribute}Mosaic plots for three affective and three instrumental values; values on the x-axis represent proportion of modes selected for each value, and values on the y-axis represent proportion of respondents by main mode of transportation"}

ggplot(data = Modes_by_Attribute %>% filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"))) + 
  geom_mosaic(aes(product(IDEAL, USED), fill = IDEAL, weight = WEIGHT),
              divider = mosaic("vertical")) + 
  xlab("Proportion of responses by mode associated with {Freedom, Enjoyment, Happiness, Poverty, Luxury, Status}") +
  ylab("Proportion of respondents by primary mode of travel") +
  scale_fill_viridis_d(direction = -1) +
  facet_wrap(~ ATTRIBUTE) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90))
```

### 2019-10-23 Dissonance_Santiago_v1.tex
Commit files from Paez-Laptop

  \title{Do Drivers Dream of Walking? An Investigation of Dissonance in the
Primary Mode of Travel from the Perspective of Subjective Wellbeing}
    \author[Some University]{Author 1\corref{c1}}
   \ead{a1@university.com} 
   \cortext[c1]{Corresponding Author}
    \author[Some School]{Author 2}
   \ead{a2@school.com}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-attribute}Mosaic plots for affective
values; the y-axis is the proportion of respondents by main mode of
transportation, and the x-axis is the proportion of modes selected for
each value}
\end{figure}

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-age-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-age}Mosaic plots for affective
values by age; the y-axis is the proportion of respondents by main mode
of transportation, and the x-axis is the proportion of modes selected
for each value}
\end{figure}

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-education-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-education}Mosaic plots for affective
values by education; the y-axis is the proportion of respondents by main
mode of transportation, and the x-axis is the proportion of modes
selected for each value}
\end{figure}

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-income-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-income}Mosaic plots for affective
values by income; the y-axis is the proportion of respondents by main
mode of transportation, and the x-axis is the proportion of modes
selected for each value}
\end{figure}

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-time-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-travel-time}Mosaic plots for
affective values by travel time; the y-axis is the proportion of
respondents by main mode of transportation, and the x-axis is the
proportion of modes selected for each value}
\end{figure}

>>>>

  \title{Do Drivers Dream of Walking? An Investigation of Dissonance in the
Primary Mode of Travel from the Perspective of Subjective Wellbeing}
    \author[Some University]{Beasly Bee\corref{c1}}
   \ead{alice@university.com} 
   \cortext[c1]{Corresponding Author}
    \author[Some School]{Antsy Aardvark}
   \ead{aa@school.com}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-without-instrumental-1.pdf}
\caption{\label{fig:mosaic-plots-by-attribute}Mosaic plots for three
affective and three instrumental values; values on the x-axis represent
proportion of modes selected for each value, and values on the y-axis
represent proportion of respondents by main mode of transportation}
\end{figure}
